Insights Into Computational Reproducibility in Plant Pathology and a Way Forward

### 2021-02-25 analysis/paper/paper.Rmd
Fix the caption for Figure 5

> "The closeness of agreement between independent results obtained with the same method on identical test material, under the same conditions (same operator, same apparatus, same laboratory and after short intervals of time). The measure of repeatability is the standard deviation qualified with the term: `repeatability' as repeatability standard deviation. In some contexts repeatability may be defined as the value below which the absolute difference between two single test results obtained under the above conditions, may be expected to lie with a specified probability", [@McNaught1997].

@@

However, while these definitions are clear and well accepted in physical sciences, in biological sciences it is not always possible to use identical test material or perhaps the the time or resources are not available for full reproduction, _e.g._, field trials that span years and locations or complex glasshouse experiments.
Therefore we will follow Peng's [-@Peng2009] definition that provides clear guidelines a minimum standard of "reproducible research":

> "The replication of scientific findings using independent investigators, methods, data, equipment, and protocols has long been, and will continue to be, the standard by which scientific claims are evaluated. However, in many fields of study there are examples of scientific investigations that cannot be fully replicated because of a lack of time or resources. In such a situation, there is a need for a minimum standard that can fill the void between full replication and nothing. One candidate for this minimum standard is 'reproducible research', which requires that data sets and computer code be made available to others for verifying published results and conducting alternative analyses". Peng, R. [-@Peng2009]. _Reproducible research and Biostatistics_. Biostatistics, 10 (3): 405-408.

@@

We randomly selected 300 articles published from 2012 to 2018 across 21 journals (Table \@ref(tab:journals)) and assigned them scores reflecting their openness and reproducibility.
The journals represented both fundamental and applied areas as well as pathogen or other group-specific publications.
A list of randomly selected pages was assigned to a randomized list of the 21 journals where the page number fell within an article for the given journal.
In cases where an article was not suitable, _e.g._, a review or otherwise not related to plant pathology, the next article was selected until a suitable article was found.
Notes regarding the selection of articles can be found in the file, XXXX, available in this paper's repository.
The pages list was numbered from page one and went to page 150.
This was done since some journals restart their numbering with each issue and also ensures that the journal is more likely to have a page number corresponding to the randomly generated value.
This also assumes that there is no effect or bias on reproducibility based on the time of year that an article was published since most journals start with page number one at the beginning of the year.
The list of journals was saved from Google's sheets as a comma-separated-value (CSV) file for later use and distribution in a research compendium.

@@

  * **computational methods available**

    * 0 - Not available or not mentioned in the publication

    * 1 - Uses expensive proprietary software that only institutions would typically purchase (_i.e._, 
    ArcGIS standard is $7000)

    * 2 - Uses proprietary software that most individuals can afford (_e.g._, Excel, Matlab? ($500))

    * 3 - Uses entirely open source and free software (_e.g._, R, Julia, Python)

    * NA - No software was used in the research that can be determined as the article is written

  * **software available**

    * 0 - Not available or not mentioned in the publication

    * 1 - Uses expensive proprietary software that only institutions would typically purchase (_i.e._, 
    ArcGIS standard is $7000US)

    * 2 - Uses proprietary software that most individuals can afford (_e.g._, Excel) or SAS that is now available free for use with a university login

    * 3 - Uses entirely open source and free software (_e.g._, R, Julia, Python).

    * NA - No software was used in the research that can be determined as the article is written

  * **software cited**

    * 0 - not mentioned

    * 1 - Software mentioned by name only

    * 2 - Software cited with version number

    * 3 - All software components (SAS PROCs, R, Julia or Python packages) etc., cited

    * NA - No software was used in the research that can be determined as the article is written.

  * **data available**

    * 0 - Not available or not mentioned in the publication

    * 1 - Available upon request to the author

    * 2 - Online, but inconvenient or non-permanent (_e.g._, login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)

    * 3 - Freely available online to anonymous users for foreseeable future (_e.g._, archived using Zenodo, dataverse or university library or some other proper archiving system)

@@

A total reproducibility score was calculated as the per cent of total possible for a given paper,
_e.g._ some papers did not use specialized software therefore they were not scored for this category.

@@

At the second level, Lehner et al. [-@Lehner2017] used GitHub to host a code repository of their research compendium website with data and a reproducible report that explains in detail all steps of the analysis and the R code for conducting a meta-analysis for assessing heterogeneity in the relationship between white mold incidence and soybean yield and between incidence and soybean tied (https://emdelponte.github.io/paper-white-mold-meta-analysis/).
The website clearly demonstrates the analysis for readers and uses R so that anyone can easily replicate the work.
Likewise, Duku et al. [-@Duku2016] provide scripts for models, data and code for graphs, via a website hosted by GitHub, http://adamhsparks.github.io/MICCORDEA/, that can be used to replicate their study modelling the effects of climate change on rice bacterial blight and rice leaf blast in Tanzania.

@@

The fourth level includes packaging the analysis as a software package, _e.g._ R Julia or Python package at a minimum, with the use of continuous integration and possibly a software container like Docker as Khaliq et al. [@Khaliq2020] have done. Their research compendium provides a Docker instance, DOI and full R package that enable readers to fully replicate their analysis of chickpea ascochyta conida dispersal using the data collected and stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].

@@

* Example from Grünwald lab: 
  - paper http://apsjournals.apsnet.org/doi/full/10.1094/PHYTO-12-14-0350-FI
  - github repo https://github.com/grunwaldlab/Sudden_Oak_Death_in_Oregon_Forests

@@

When making your science more open and reproducible, methods, software used, this includes items such as R, Julia or Python packages that were directly used in the analysis or production of the paper, etc. should be cited properly.
Deposit and annotate biological materials with an herbarium or other repository.
Provide a full description of all equipment used, _e.g._ a Spectrum Technologies Watchdog 2700 weather station was used to record wind speed, direction, rainfall, temperature and relative humidity at one-hour intervals.
This allows end-users to identify what was used and identify the methods used more accurately.
Just as importantly, this acknowledges the contributions of others whose works were instrumental in your research.
This also helps ensure that you as a researcher can reconstruct what you have done since you will have good notes and documentation and be able to identify if something changes, _e.g._ a package version, what effect it had on your research.

@@

Prefer text files for saving small sets of data.
Data that are saved in binary formats such as PDF files are difficult to reuse because they are not easily machine readable.
In many cases, data sets are small enough and curated in spreadsheets, which should be saved as a plain text file, _e.g._ comma separated (CSV) files.
This also helps ensure that the data are reusable.
Larger data sets may warrant the use of a proper database like MariaDB or PostgreSQL which provide users with several benefits but two important ones to mention here are data redundancy, no duplicated records; and data consistency, all of your records are sure to be recorded in the same way for every observation.
While databases may offer many advantages, the trade-off is that they are more complex to set up and administer, especially for a small data set that might be only a few rows and columns in a database.

@@

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

@@

```{r software_used, tab.id="su", tab.cap="ally of reproducibility scores where 0 is not reproducible and 12 is fully reproducible, _e.g._ all software cited, software used are freely available to anyone, computational methods (statistical analysis or other computational code) are freely available for download and use, and raw data are available for download and use.", echo=FALSE, message=FALSE, warning=FALSE}
library("tidyr")

>>>>

> "The closeness of agreement between independent results obtained with the same method on identical test material, under the same conditions (same operator, same apparatus, same laboratory and after short intervals of time). The measure of repeatability is the standard deviation qualified with the term: \`repeatability' as repeatability standard deviation. In some contexts repeatability may be defined as the value below which the absolute difference between two single test results obtained under the above conditions, may be expected to lie with a specified probability", [@McNaught1997].

@@

However, while these definitions are clear and well accepted in physical sciences, in biological sciences it is not always possible to use identical test material or perhaps the the time or resources are not available for full reproduction, *e.g.*, field trials that span years and locations or complex glasshouse experiments.
Therefore we will follow Peng's [-@Peng2009] definition that provides clear guidelines a minimum standard of "reproducible research":

> "The replication of scientific findings using independent investigators, methods, data, equipment, and protocols has long been, and will continue to be, the standard by which scientific claims are evaluated. However, in many fields of study there are examples of scientific investigations that cannot be fully replicated because of a lack of time or resources. In such a situation, there is a need for a minimum standard that can fill the void between full replication and nothing. One candidate for this minimum standard is 'reproducible research', which requires that data sets and computer code be made available to others for verifying published results and conducting alternative analyses".
> Peng, R.
> [-@Peng2009].
> *Reproducible research and Biostatistics*.
> Biostatistics, 10 (3): 405-408.

@@

We randomly selected 300 articles published from 2012 to 2018 across 21 journals (Table \@ref(tab:journals)) and assigned them scores reflecting their openness and reproducibility.
The journals represented both fundamental and applied areas as well as pathogen or other group-specific publications.
A list of randomly selected pages was assigned to a randomized list of the 21 journals where the page number fell within an article for the given journal.
In cases where an article was not suitable, *e.g.*, a review or otherwise not related to plant pathology, the next article was selected until a suitable article was found.
Notes regarding the selection of articles can be found in the file, XXXX, available in this paper's repository.
The pages list was numbered from page one and went to page 150.
This was done since some journals restart their numbering with each issue and also ensures that the journal is more likely to have a page number corresponding to the randomly generated value.
This also assumes that there is no effect or bias on reproducibility based on the time of year that an article was published since most journals start with page number one at the beginning of the year.
The list of journals was saved from Google's sheets as a comma-separated-value (CSV) file for later use and distribution in a research compendium.

@@

-   **computational methods available**

    -   0 - Not available or not mentioned in the publication

    -   1 - Uses expensive proprietary software that only institutions would typically purchase (*i.e.*, ArcGIS standard is \$7000)

    -   2 - Uses proprietary software that most individuals can afford (*e.g.*, Excel, Matlab? (\$500))

    -   3 - Uses entirely open source and free software (*e.g.*, R, Julia, Python)

    -   NA - No software was used in the research that can be determined as the article is written

-   **software available**

    -   0 - Not available or not mentioned in the publication

    -   1 - Uses expensive proprietary software that only institutions would typically purchase (*i.e.*, ArcGIS standard is \$7000US)

    -   2 - Uses proprietary software that most individuals can afford (*e.g.*, Excel) or SAS that is now available free for use with a university login

    -   3 - Uses entirely open source and free software (*e.g.*, R, Julia, Python).

    -   NA - No software was used in the research that can be determined as the article is written

-   **software cited**

    -   0 - not mentioned

    -   1 - Software mentioned by name only

    -   2 - Software cited with version number

    -   3 - All software components (SAS PROCs, R, Julia or Python packages) etc., cited

    -   NA - No software was used in the research that can be determined as the article is written.

-   **data available**

    -   0 - Not available or not mentioned in the publication

    -   1 - Available upon request to the author

    -   2 - Online, but inconvenient or non-permanent (*e.g.*, login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)

    -   3 - Freely available online to anonymous users for foreseeable future (*e.g.*, archived using Zenodo, dataverse or university library or some other proper archiving system)

@@

A total reproducibility score was calculated as the per cent of total possible for a given paper, *e.g.* some papers did not use specialized software therefore they were not scored for this category.

@@

At the second level, Lehner et al. [-@Lehner2017] used GitHub to host a code repository of their research compendium website with data and a reproducible report that explains in detail all steps of the analysis and the R code for conducting a meta-analysis for assessing heterogeneity in the relationship between white mold incidence and soybean yield and between incidence and soybean tied (<https://emdelponte.github.io/paper-white-mold-meta-analysis/>).
The website clearly demonstrates the analysis for readers and uses R so that anyone can easily replicate the work.
Likewise, Duku et al. [-@Duku2016] provide scripts for models, data and code for graphs, via a website hosted by GitHub, <http://adamhsparks.github.io/MICCORDEA/>, that can be used to replicate their study modelling the effects of climate change on rice bacterial blight and rice leaf blast in Tanzania.

@@

The fourth level includes packaging the analysis as a software package, *e.g.* R Julia or Python package at a minimum, with the use of continuous integration and possibly a software container like Docker as Khaliq et al. [@Khaliq2020] have done.
Their research compendium provides a Docker instance, DOI and full R package that enable readers to fully replicate their analysis of chickpea ascochyta conida dispersal using the data collected and stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].

@@

-   Example from Grünwald lab:

    -   paper <http://apsjournals.apsnet.org/doi/full/10.1094/PHYTO-12-14-0350-FI>
    -   github repo <https://github.com/grunwaldlab/Sudden_Oak_Death_in_Oregon_Forests>

@@

When making your science more open and reproducible, methods, software used, this includes items such as R, Julia or Python packages that were directly used in the analysis or production of the paper, etc. should be cited properly.
Deposit and annotate biological materials with an herbarium or other repository.
Provide a full description of all equipment used, *e.g.* a Spectrum Technologies Watchdog 2700 weather station was used to record wind speed, direction, rainfall, temperature and relative humidity at one-hour intervals.
This allows end-users to identify what was used and identify the methods used more accurately.
Just as importantly, this acknowledges the contributions of others whose works were instrumental in your research.
This also helps ensure that you as a researcher can reconstruct what you have done since you will have good notes and documentation and be able to identify if something changes, *e.g.* a package version, what effect it had on your research.

@@

Prefer text files for saving small sets of data.
Data that are saved in binary formats such as PDF files are difficult to reuse because they are not easily machine readable.
In many cases, data sets are small enough and curated in spreadsheets, which should be saved as a plain text file, *e.g.* comma separated (CSV) files.
This also helps ensure that the data are reusable.
Larger data sets may warrant the use of a proper database like MariaDB or PostgreSQL which provide users with several benefits but two important ones to mention here are data redundancy, no duplicated records; and data consistency, all of your records are sure to be recorded in the same way for every observation.
While databases may offer many advantages, the trade-off is that they are more complex to set up and administer, especially for a small data set that might be only a few rows and columns in a database.

@@

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

@@

```{r software_used, tab.id="su", tab.cap="Tally of reproducibility scores where 0 is not reproducible and 12 is fully reproducible, _e.g._ all software cited, software used are freely available to anyone, computational methods (statistical analysis or other computational code) are freely available for download and use, and raw data are available for download and use.", echo=FALSE, message=FALSE, warning=FALSE}
library("tidyr")
